---
title: "PA1_template"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###Loading and preprocessing the data


```{r }
dataset_url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip" 
download.file(dataset_url, "Activity.zip")
unzip("Activity.zip", exdir = "Activity")

Act <- data.table::fread("Activity/activity.csv") 
Act[, date := as.POSIXct(date, format = "%Y-%m-%d")] #Change character tot  to date
str(Act)

```

### Calculate the total number of steps taken per day

```{r}
Day_total <- aggregate(steps ~ date, data = Act, sum, na.rm = TRUE)
with(Day_total,plot(date,steps,
              type = "h",
              lwd = 4,
              col="red",
              main="Total steps per day",
              xlab="", 
              ylab="Steps"))
```


```{r}
mean <- as.integer(mean(Day_total$steps))
median <- median(Day_total$steps)
```

The mean is `r mean` and the median is `r median` steps per day

### Make a time series plot (i.e. type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)


```{r}
Act_CC <- Act[complete.cases(Act), ]

Interval_mean = aggregate(steps ~ interval, data = Act_CC, FUN = "mean", na.rm = TRUE)


with(Interval_mean,plot(interval,steps,
                        type = "l",
                        lwd = 4,
                        col="red",
                        main="Avg steps per 5 min interval",
                        xlab="Interval", 
                        ylab="Steps"))
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max_steps <- Interval_mean$interval[max.col(t(Interval_mean$steps))] 
```

The max number of steps is found in the interval: `r max_steps` 

### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
NA_Count <- Act[is.na(steps), .N ] #Number of missing values
Total_Rows <- Act[, .N ] #Number of rows
NA_Percentage <- round((NA_Count / Total_Rows) *100,2)

```

The number of missing values is `r NA_Count` (`r NA_Percentage`%)

### Create a new dataset that is equal to the original dataset but with the missing data filled in.

I decided to replace the missing values with te avarage number of steps of that interval
```{r}
Act_NA <- data.frame(Act)

library(zoo)
Act_NA <- transform(Act_NA, steps = na.aggregate(steps, by = interval)) 
summary(Act_NA)
```

### Make a histogram of the total number of steps taken each day 

```{r}
Day_total_NA <- aggregate(steps ~ date, data = Act_NA, sum)

with(Day_total_NA,plot(date,steps,
                    type = "h",
                    lwd = 4,
                    col="red",
                    main="Total steps per day",
                    xlab="", 
                    ylab="Steps"))
```

### Calculate and report the mean and median total number of steps taken per day. 

```{r}
Mean_after <- as.integer(mean(Day_total_NA$steps))
Median_after <- as.integer(median(Day_total_NA$steps))
```

The mean after handling the missing values is `r Mean_after` and the median is `r Median_after` 

### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends


```



```{r}
Sys.setlocale("LC_TIME", "en_US") #Set sys language to English


Act_NA$Weekday <- as.factor(weekdays(Act_NA$date))
Act_NA$Weekend_ind = as.factor(ifelse(Act_NA$Weekday == "Saturday" | Act_NA$Weekday == "Sunday" ,"Weekend","Weekday"))

Weekend_interval <- aggregate(steps ~  interval + Weekend_ind, data = Act_NA, FUN = "mean")
Weekend_interval$steps <- as.integer(Weekend_interval$steps)

library(ggplot2)

ggplot(Weekend_interval, aes(interval, steps)) +
  geom_line(stat = "identity", aes(colour = Weekend_ind)) +
  theme_gray() +
  facet_grid(Weekend_ind ~ ., scales="fixed", space="fixed") +
  labs(x="Interval", y=expression("Steps")) +
  ggtitle("Weekday versus Weekend days")

```


